---
output: 
  github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

glmtools
=====
[![Build status](https://ci.appveyor.com/api/projects/status/j5mscylmkssa0esf)](https://ci.appveyor.com/project/jread-usgs/glmtools) 
[![Build Status](https://travis-ci.org/USGS-R/glmtools.svg)](https://travis-ci.org/USGS-R/glmtools)
<!-- [![Coverage Status](https://img.shields.io/coveralls/USGS-R/glmtools.svg)](https://coveralls.io/r/USGS-R/glmtools) -->

Tools for interacting with the [General Lake Model (GLM)](http://aed.see.uwa.edu.au/research/models/GLM/ "General Lake Model's website") in R. `glmtools` includes some basic functions for calculating physical derivatives and thermal properties of model output, and some plotting functionality (see example image below). 

`glmtools` Functions (as of v0.2.5.2)
=====
```{r load printr, echo=FALSE,message=FALSE,results='hide'}
loadNamespace("printr")
```

```{r echo=FALSE}
help.search("*", package = "glmtools", types = "help")
```

```{r unload printr, echo=FALSE}
unloadNamespace("printr")
```

FAQ
=====

How do I install `glmtools`? It isn't on CRAN!
===
```r
install.packages('remotes')
remotes::install_github('usgs-r/glmtools')
```
        
What libraries does `glmtools` need?
===
This version requires the NetCDF version R library (called `ncdf4`), and `rLakeAnalyzer`, and `tools` if you would like to run all examples. 

```{r echo=FALSE}
deps <- desc::desc_get_deps()
knitr::kable(deps[deps$type %in% c("Imports", "Depends"), 2:3])
```

How do I contribute new code back to the glmtools project?
===

In order to contribute to this code, we recommend the following workflow: 

1) "fork" this repository to your own personal github account

2) clone the github repository to your computer:

	$git clone https://github.com/{username}/glmtools.git

3) modify code or add new functionality, save the code

4) add the repository master to a remote master called "upstream"

	$cd glmtools

	$git remote add upstream https://github.com/USGS-R/glmtools.git

5) before pushing your changes to your repository, pull in the current version of the GLEON master:

	$git fetch upstream

6) merge these differences with your own "master" version:

	$git merge upstream/master

7) push your changes to your github repository, in addition to changes made by pulling in the GLEON master:

	$git push

8) submit a pull request to GLEON master using your account at github.com

