---
title: "Walkthrough of the calibration approach"
author: "Jordan S Read, Luke A Winslow, Hilary A Dugan, Robert Ladwig, Tadhg Moore"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Walkthrough of the calibration approach}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
library(rmarkdown)
options(continue=" ")
options(width=60)
library(knitr)
library(glmtools)
knitr::opts_knit$set(root.dir = tempdir())
setwd(tempdir())
temp_folder <- tempdir()
out_file <- file.path(temp_folder, 'output/output.nc')
field_data <- file.path(temp_folder, 'LakeMendota_field_data_hours.csv')
file.copy('glm3_uncalibrated.nml', 'glm3.nml', overwrite = TRUE)
nml_data <- file.path(temp_folder, 'glm3.nml')
```


## Introduction




## Using example data built into the package

### Find example data directory
```{r}
library(glmtools)
GLM3r_folder = system.file('extdata', package = 'GLM3r')
glmtools_folder = system.file('extdata', package = 'glmtools')
```
### List files that are included in the example data
```{r}
dir(GLM3r_folder)
dir(glmtools_folder)
```

### Move glmtools example files to a temporary folder
```{r}
example.files <- dir(glmtools_folder)
temp_folder <- tempdir()

for(file in paste0(glmtools_folder,'/',example.files)) {
  file.copy(file, temp_folder, overwrite = TRUE)
}
```

Note that new output files have been added to the temporary directory:
```{r}
dir(temp_folder)
```

### Visualizing the uncalibrated setup
We can use `glmtools` functions to visualize and quantify the goodness of the fit of the uncalibrated setup, for instance by using the plotting and comparing functions when field data is available. But first, we need to declare the names of our output, configuration and field data variables.
```{r}
out_file <- file.path(temp_folder, 'output/output.nc')
field_data <- file.path(temp_folder, 'LakeMendota_field_data_hours.csv')
file.copy('glm3_uncalibrated.nml', 'glm3.nml', overwrite = TRUE)
nml_data <- file.path(temp_folder, 'glm3.nml')
```
## Running uncalibrated GLM setup
The contents of `glmtools_folder` (i.e., a `glm3.nml` file and driver data including `.csv` file for the meteorological drivers) are now located in a temporary folder (`temp_folder`). The `run_glm()` function will now return the results from the uncalibrated setup.
```{r, results ='hide'}
run_glm(sim_folder = temp_folder)
```

Then we can plot the uncalibrated setup and quantify the root-mean squared error between observed and simulated data.

```{r}
plot_var_nc(out_file, var_name = 'temp')
plot_var_compare(nc_file = out_file, field_file = field_data,var_name = 'temp', precision = 'hours')

temp_rmse <- compare_to_field(out_file, field_file = field_data,
                              metric = 'water.temperature', as_value = FALSE, precision= 'hours')
print(paste('total time period:',round(temp_rmse,2),'deg C RMSE'))
```

```{r}
var = 'temp'         # variable to which we apply the calibration procedure
path = getwd()       # simulation path/folder
obs = read_field_obs(field_data)  # observed field data
nml.file = nml_data  # path of the nml configuration file
calib_setup = get_calib_setup() # create a setup of variables for calibration
glmcmd = NULL        # command to be used, default applies the GLM3r function
# Optional variables
first.attempt = TRUE # if TRUE, deletes all local csv-files that stores the outcome of previous calibration runs
period = get_calib_periods(nml = nml_data, ratio = 1) # define a period for the calibration, thissupports a split-sample calibration (e.g. calibration and validation period)
scaling = TRUE       # scaling of the variables in a space of [0,10]; should be TRUE for CMA-ES
method = 'CMA-ES'    # optimization method, choose either `CMA-ES` or `Nelder-Mead`
metric = 'RMSE'      # objective function, which will be minimized, here the root-mean square error
target.fit = 1.0     # refers to a target fit of 1.0 degrees Celsius
target.iter = 50    # refers to a maximum run of 150 calibration iterations
plotting = FALSE      # if set to TRUE, the script will automatically save the created contour plots 
output = out_file    # path of the output file
field.file = field_data # path of the field data

# main calibration function
calibrate_sim(var, path, obs, nml.file, calib_setup, glmcmd, first.attempt, period, scaling, method, metric, target.fit, target.iter, plotting, output, field.file)

```